\chapter{Introduction}
\label{ch:introduccion}

The Cosmos SDK is an open source blockchain development framework that facilitates the development of tailor made blockchain applications. The Tendermint consensus mechanism, a quick and safe proof-of-stake consensus method, is the foundation upon which the Cosmos SDK is constructed. As a result, developers are freed from having to worry about the supporting infrastructure and can concentrate on creating the application layer as we will see in Chapter \ref{OCS:overview}.

Private blockchain adoption has grown in popularity in recent years as businesses and organisations look to capitalize on the advantages of blockchain technology without the constraints of public blockchains. Comparatively speaking, private blockchains provide far more efficiency, privacy, and control. With its modular design and adaptability, the Cosmos SDK provides an ideal environment for developing private blockchain solutions.

The technical features of the Cosmos SDK and the Tendermint consensus mechanism, as well as their advantages and drawbacks will be examined in this thesis. The Cosmos SDK's major applications for private blockchains will also be explored at. Finally, to demonstrate how to use this framework in real world applications, we will provide a case study of a private blockchain application created with the Cosmos SDK.

\section{Background and Motivation}

Blockchain technology has gained significant attention in recent years for its potential to disrupt various industries and change the world for the better. As a result, I started working on projects like \glspl{nft}, smart contracts, and \gls{defi} protocols in the cryptocurrency business. The potential of blockchain technology, however, stretches far beyond these applications.

This technology has the power to empower individuals by removing intermediaries and enabling \gls{p2p} transactions. It can reduce transaction costs, increase efficiency, and promote greater financial inclusion, particularly in developing countries where access to financial services is limited, and corruption is rampant.

Motivated by its potential to create positive social impact, the thesis aims explore the transformative potential of the Cosmos SDK and the Tendermint consensus mechanism for private blockchains. By examining the technical aspects and exploring possible applications for private blockchains, contributing to the growing body of research on the transformative potential of blockchain technology.

\section{Document structure}

This paper is structured as follows:

\begin{itemize}
    \item Section \ref{OCS:overview} provides deeper insight into the Cosmos SDK its architecture and how developers leverage the framework.
    \item Section \ref{ch:compare} provides a comparison of different consensus algorithms and their applicability for different use cases.
    \item Section \ref{ch:private} provides a description of the design process of a private blockchain with a real application case.
    \item Section \ref{ch:conclusion} provides  a conclusion of the concepts covered during the thesis, as well as different paths for future work.
\end{itemize}

